---
title: "Function Development"
author: "Josephine Ward"
date: "2025-02-10"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# TOC
* I. Packages
* II. Transition Matrix from an Individual Focal
* III. New Function: Get Log-Likelihood
  - III.1 Pseudo-Code
* IV. New Function: Get Behaviors
  - IV.1 Pseudo-Code

# I. Packages

```{r}

```


# II. Transition Matrix from an Individual Focal

```{r}
behaviorshortlist <-c("Move", "Vigilance", "Feed", "Rest",
               "Affiliative", "Aggressive", "SelfGroom",
               "Survey", "PAUSE", "Human Directed",
               "Forage", "Sleep", "Play")

matrixfunct <- function(dataframe, behaviorshortlist){
  Matrix <- matrix(NA, nrow = length(behaviorshortlist), ncol = length(behaviorshortlist))
  rownames(Matrix) <- colnames(Matrix) <- behaviorshortlist
  for(b1 in behaviorshortlist){
    denomentator <- sum(dataframe$TimeSpent[dataframe$Behavior == b1]) - 
      as.numeric(tail(dataframe$Behavior, 1) == b1) ## accounting for if the last behavior in focal is b1 (doesn't count as a transition opportunity), overall total possible transition opportunities
    b1_ind <- which(dataframe$Behavior == b1)
    for(b2 in behaviorshortlist){
      if(b1 == b2){ ## if staying in the same behavior
        numerator <- denomentator - sum(dataframe$Behavior == b1) + 
          as.numeric(tail(dataframe$Behavior, 1) == b1) 
        ## removing the times they switched to a diff behavior & adding back in that situation if the last behavior is b1
        } else { ## if moving to a different behavior
          numerator <- sum(dataframe$Behavior[b1_ind + 1] == b2, na.rm = T)
          }
      Matrix[b1,b2] <- numerator/denomentator
    }
  }
return(Matrix)
}

Matrix <- matrixfunct(focal28, behaviorshortlist)
rowSums(Matrix)
Matrix
```

** NOTE: this does not work when we have the situation where the data file has 2 of the same behavior right on top of eachother --> probably easier to fix by cleaning the dataframe than changing the function (also we never concretely established why it is like that in the data)


** future consideration: how to make this apply to multiple combined focals
** also make it more generalizable than the specific names I have going on here


# III. Get Log-Likelihood Function

## III.1 Pseudo-Code/Work
- get log-likelihood function (the spacing on this will probably look wack when i knit but thats life)
_______________________________________________________
Outline from Henry:
P = the probability transition matrix?

get_LL --> function(behaviors, P){
  .
  .
  .
  return(log-likelihood)
}
Check:
If you’ve made behaviors w/particular P, does it give you close to the highest possible value with the first function
1. Simulate behaviors from P
2. Check get_LL is maximized close to P
Then optim function will help us with maximization

_________________________________________________________
ok idk what im doing here
ive potentially lost the plot with what this function is supposed to accomplish 

so we want the likelhood --> probability of being in any particular behavior
so this function needs to take in data and calculate probability of being in each of the behaviors
that is not matching up with what henry gave above
is the output a vector? like a scalar value for each behavior and then the entire vector adds to 1?

are we assuming we have enough data to make this a normal distribution? is that a bold assumption it feels bold but then we can use dnorm/rnorm???

optim() function for optimization (so like do the derivative part for me once i have the equation?)

________________________________________________________________________
get_LL --> function(list of behaviors, data){
  n <- length(behaviors) ## number of observations? idk if behaviors is what to put there
  P <- matrixfunct(data, behaviors) ## so this will give us the transition matrix, only works for 1 focal rn
  
  liklihood_eq <- (product (of each entry of the matrix^total number of transitions?))
    ## no thats not right bc it will just give a single number? 
    ## do i do this for each behavior? should this be in the for loop
 
  log-likelihood <- define an empty matrix here with length(behaviors) to add to in the for loop
  
  for(b in behaviors){
    optim(log(liklihood_eq))
    and then add it to the matrix in the right spot in log-likelihood
  }
  return(log-likelihood)
}
___________________________________________________________________________

# IV. Get Behaviors

## IV.1 Pseudo-Code

from henry:

get_behaviors —> function(P, initial_behavior){
	.
	.
	.
	return(Behaviors)}
