---
title: "Function Development"
author: "Josephine Ward"
date: "2025-02-10"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# TOC
* I. Packages
* II. Transition Matrix from an Individual Focal
* III. New Function: Get Log-Likelihood
  - III.1 Pseudo-Code
  - III.2 Code Attempts/notes idk
* IV. Isolate Focal Function
* V. Optimizing LL

# I. Packages

```{r}
library(dplyr) ## need for case_when in get_ll function
```

## data import

```{r}
monkeyprelimdata <- read.csv("~/Desktop/MonkeyProject/monkeyprelimdata.csv")
```

# II. Transition Matrix from an Individual Focal

```{r}
behaviorshortlist <-c("Move", "Vigilance", "Feed", "Rest",
               "Affiliative", "Aggressive", "SelfGroom",
               "Survey", "Human Directed",
               "Forage", "Sleep", "Play")

matrixfunct <- function(dataframe, behaviorshortlist){
  Matrix <- matrix(NA, nrow = length(behaviorshortlist), ncol = length(behaviorshortlist))
  rownames(Matrix) <- colnames(Matrix) <- behaviorshortlist
  for(b1 in behaviorshortlist){
    denomentator <- sum(dataframe$TimeSpent[dataframe$Behavior == b1]) - 
      as.numeric(tail(dataframe$Behavior, 1) == b1) ## accounting for if the last behavior in focal is b1 (doesn't count as a transition opportunity), overall total possible transition opportunities
    b1_ind <- which(dataframe$Behavior == b1)
    for(b2 in behaviorshortlist){
      if(b1 == b2){ ## if staying in the same behavior
        numerator <- denomentator - sum(dataframe$Behavior == b1) + 
          as.numeric(tail(dataframe$Behavior, 1) == b1) 
        ## removing the times they switched to a diff behavior & adding back in that situation if the last behavior is b1
        } else { ## if moving to a different behavior
          numerator <- sum(dataframe$Behavior[b1_ind + 1] == b2, na.rm = T)
          }
      Matrix[b1,b2] <- numerator/denomentator
    }
  }
return(Matrix)
}
```

** NOTE: this does not work when we have the situation where the data file has 2 of the same behavior right on top of eachother --> probably easier to fix by cleaning the dataframe than changing the function (also we never concretely established why it is like that in the data)


** future consideration: how to make this apply to multiple combined focals
** also make it more generalizable than the specific names I have going on here


# III. Get Log-Likelihood Function

## III.1 Pseudo-Code/Work
- get log-likelihood function (the spacing on this will probably look wack when i knit but thats life)
_______________________________________________________
Outline from Henry:
P = the probability transition matrix?

get_LL --> function(behaviors, P){
  .
  .
  .
  return(log-likelihood)
}
Check:
If you’ve made behaviors w/particular P, does it give you close to the highest possible value with the first function
1. Simulate behaviors from P
2. Check get_LL is maximized close to P
Then optim function will help us with maximization

___________________________________________________________________________
work w/Henry during the meeting:

y1 = behavior at first time point
y1| p1 ~ categorical w/probability p1= (p11, p12, ..., p1B) B = 13/ whatever the number of behaviors are
  p1 = probabilkty of being in which of the B behaviors when you wander out and see the monkeys

P = transition matrix 

yt = ( 0 0 0 ...... 1 ...... 0 0 0)^T bth behavior

yt|[y(t-1), P ] ~ cat(Py(t-1)) ~ Cat(pt)

pt = P*y(t-1) --> picks out particular row of P

holds for t2 --> B

so now we can find the likelihood
things:
- categorical --> pmf not pdf
- not independent 

L(theta, y) = P(y1|p1)P(y2|y1, P)P(y3|y2, P) .....
            = P(y1|p1)*product(P(yt|y(t-1), P))
            
            inner product of y1^t*p1
            (y1^t*p1)(yT*P*y(T-1)).....(y2*P*y(2-1))
            
we prefer log

log(L) = sum of logs = log(y1^t*p1) + log(yT*P*y(T-1)) + ......

P(y1|p1) = P11 ^ y11 * P12^y12 * ..... * P1B^y1B
log(^^) = y1^tlog(p1)

need to like expand the data so it has each second as individual 

output is a single scalar (not a matrix)

essentially doing this to test different Ps

then we optimize that function to find the Ps


test function , give it a P, tells us how good we are doing

checks we can do:

take a focal, take one of the point estimated matrices, do the log-likelihood, try it again with a diff transition matrix, likelihood value should go down

P = matrix(p11, p12, p13, p21, p22, p23, p31, p32, p33, row = 3, col = 3)
but what if p12 = beta0 + beta1(X=indicator for forest)
probably have to do logit and make sure the ps all sum to 1 still? idk 
so we take our function and tweak it to have betas instead of Ps
___________________________________________________________________________
## III.2 Code Attempts
```{r}
## note that you need the same order of behaviors that I have (probs can change that later)
behaviorshortlist <-c("Move", "Vigilance", "Feed", "Rest",
               "Affiliative", "Aggressive", "SelfGroom",
               "Survey", "Human Directed",
               "Forage", "Sleep", "Play") 

## functions takes data (a focal) and P (a transition matrix)
get_ll <- function(data, P, behavior_list){
  B <- nrow(P) # number of behaviors
  y <- behavior_list
  ymatrix <- model.matrix(~y - 1)
  
  ## build a Y Duration Matrix (which is just P without the denominator I think)
  Yduration <- matrix(NA, nrow = length(behavior_list), ncol = length(behavior_list))
  rownames(Yduration) <- colnames(Yduration) <- behavior_list
  for(b1 in behavior_list){
    denomentator <- sum(data$TimeSpent[data$Behavior == b1]) - 
      as.numeric(tail(data$Behavior, 1) == b1)
    b1_ind <- which(data$Behavior == b1)
    for(b2 in behavior_list){
      if(b1 == b2){
        numerator <- denomentator - sum(data$Behavior == b1) + 
          as.numeric(tail(data$Behavior, 1) == b1) 
        } else { 
          numerator <- sum(data$Behavior[b1_ind + 1] == b2, na.rm = T)
          }
      Yduration[b1,b2] <- numerator
    }
  }
  
  LogL <- sum((log(P)*Yduration)[Yduration>0])
  
  if(is.na(LogL)){
    return(-Inf)
  } else {
    return(LogL)
  }
}
```


```{r}
## ok im gonna train it on focal 28 and then use focal 10 as the input data?

P_28 <- matrixfunct(focal_28, behaviorshortlist)
P_28
P_28[is.na(P_28)] <- 0 ## get rid of NaNs

focal_10 <- isolate.focal(10)
P_10 <- matrixfunct(focal_10, behaviorshortlist)
P_10[is.na(P_10)] <- 0 ## get rid of NaNs

get_ll(focal_10, P_28) ## so this gave -inf which makes sense but im realizing a flaw in testing bc all of the transition have like missing values (0s) so its always going to be like no???

get_ll(focal_10, P_10)

P_fake <- matrix(1/12, 12, 12) ## cant have 0 probability for a transition that we do observe
focal_1 <- isolate.focal(1)

get_ll(focal_1, P_fake)
```

#### Resolved things (that arent really resolved but im kinda just ignoring for now):

- like somewhere there is supposed to be a part where i can put some variables like forest/road or whatever and that is included in the likelihood? im not sure where to put this
  -ignore this for now (which will replace the ps with betas in the future)
-overall NaNs from P to 0s to make it work (for now, dont change the matrixfunct function)

- pause? remove but dont want the next to be a transition?
  - change og function to ignore PAUSE
  - chop off focals after point of PAUSE (lose some data)
  - split into multiple focals?
  - choose path of least resistance rn
  - for rn im just going to choose focals that dont have PAUSE and i will deal with that later :)
____________________________________________________
	
## IV. Isolate focal function
```{r}
## function to isolate a single focal & add a cumulative time column
isolate.focal <- function(i){
  onefocal.i <- subset(monkeyprelimdata, monkeyprelimdata$FocalID == i)
  cumsum.i <- data.frame(cumsum(onefocal.i$TimeSpent))
  onefocal.i <- cbind(onefocal.i, cumsum.i)
  return(onefocal.i)
}

# test
focal_28 <- isolate.focal(28)
```

# V. Optimizing LL


## Henry optim chat ----

```{r}
get_ll.h <- function(X, mu, sigma){
  if(sigma < 0) return(-Inf)
  sum(dnorm(X, mean = mu, sd = sigma, log = T))
}

get_ll.h(c(1, 2, 3, 4), mu = 2.3, sigma = 1)

fit.h <- optim(par = c(0, 5), fn = function(par, X){
  get_ll.h(X, mu = par[1], sigma = par[2])
}, X = c(1, 2, 3, 4), control = c(fnscale = -1))

fit.h

```

## applying to our function
```{r}
## fake dataset (3 behaviors only)

fake_behaviors <- c("move", "feed", "move", "sleep", "feed", "sleep", "move", "sleep", "move")
fake_duration <- c(5, 10, 4, 102, 16, 58, 14, 9, 11) 
X <- fake_data <- data.frame(fake_duration, fake_behaviors)
colnames(X) <- colnames(fake_data) <- c("TimeSpent", "Behavior") ## match my fake one to work with the matrixfunctions

fake_behaviorshortlist <- c("move", "sleep", "feed")
Pf <- matrixfunct(X, fake_behaviorshortlist) ## transition matrix of fake focal

get_ll(X, Pf, fake_behaviorshortlist) 
## note i modified the LL function to add the list of behaviors as an input so i can do mini tests/if we add or remove any behaviors

## but now i want to optimize the P (so like its unknown)
## par is the intial values (starting point to optimize from)
## fn is the function to be maximized (our logLL function)
## then we can have further arguments that get passed to fn (so that would be our X?)
## control is scaling it bc optim likes to minimize and we want to optimize

## make the optim thing within another function to make it simpler to use (note: add to package)
fit_MM <- function(X, behavior_list, par = matrix(1/length(behavior_list), 
                                                  nrow = length(behavior_list), 
                                                  ncol = length(behavior_list) - 1)){
  fit <- optim(par = par, fn = function(par, X){
    if(any(par < 0)) return(-Inf)
    if(any(par > 1)) return(-Inf)
    Pf <- matrix(par, nrow = length(behavior_list))
    if(any(rowSums(Pf) > 1)) return(-Inf)
    Pf <- cbind(Pf, 1 - rowSums(Pf))
    rownames(Pf) <- colnames(Pf) <- behavior_list
    get_ll(X, Pf, behavior_list)}, X = X, control = c(fnscale = -1))
  
  out <- fit$par
  out <- cbind(out, 1 - rowSums(out))
  rownames(out) <- colnames(out) <- behavior_list
  return(out)
}

fit_MM(X, fake_behaviorshortlist)
Pf

## fake initial matrix
initial_matrix_fake <- matrix(c(0.3, 0.3, 0.3, 0.3, 0.3, 0.3, 0.4, 0.4, 0.4), 
                              nrow = 3, ncol = 3)
row.names(initial_matrix_fake) <- fake_behaviorshortlist
colnames(initial_matrix_fake) <- fake_behaviorshortlist
```
