---
title: "Initial Monkey Stuff"
author: "Josephine Ward"
date: "2024-12-11"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# ToC
* I. Packages
* II. Initial Setup
  - II.1 Defining Stuff
* III. Basic Plots
  - III.1 Time Plots
  - III.2 Behavior Trajectory Plot Attempt (failed...)
* IV. Transition Matrices Attempts (lot of failure)
  - IV.1 trying packages bc i didnt want to do it myself
* V. Reset/ditched the script file for the rmd
  - V.1 Isolate Focal Function
* VI. Looking at Individual Focals
  - VI.1 Comparing individuals in forest vs road
* VII. Transition Matrix Function!
* VIII. Making New Functions
  - VIII.1 Pseudo-code mapping

# I. Packages
```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(igraph)
library(lubridate)
library(markovchain)
library(migrate)
```


# II. Initial Setup
```{r}
## setup
prelim <- read.csv("~/Desktop/monkey/SEED_Focal Continuous Preliminary Data_21 July 2024 to 31 August 2024_Timestamped_AD-1.csv")
summary(prelim)

## subset of data (variables I think I care about)
monkeysub <- data.frame(prelim$Time.Spent..Seconds., prelim$Focal.ID, prelim$Focal.Individual, 
                       prelim$Habitat, prelim$Behavior, prelim$Time)
labels <- c("Time", "FocalID", "Individual", "Habitat", "Behavior", "TimeStamp")
colnames(monkeysub) <- labels
monkeysub <- monkeysub[!(monkeysub$Behavior %in% "End Focal (NEXT)"),]
```

## II.1 Defining Stuff
```{r}
monkeynames <- c("Abu", "Afwan", "Beti", "Caca", "Cri", "Daffa", "Dirgha", "Erin", "Gado", "Hantu",
                 "Jaya", "Jengo", "Korra", "Mamoa", "Misses", "Moka", "Nami", "Nopi",
                 "OG", "Pado", "Rusak")

habitats <- c("Jalan", "Hutan")

## i think i renamed this later to get rid of the SAVE and NEXT stuff
behaviors <- c("Move (SAVE)", "Vigilance (SAVE)", "Feed (NEXT)", "Rest (SAVE)",
               "Affiliative (NEXT)", "Aggressive (NEXT)", "SelfGroom (SAVE)",
               "Survey (SAVE)", "PAUSE (SAVE)", "Human Directed (NEXT)",
               "Forage (NEXT)", "Sleep (SAVE)", "Play (SAVE)")
```

# III. Basic Plots

## III.1 Time Plots
```{r}
## BASIC TIME PLOTS

## plot for individual vs sum time
IndivSumTimes <- c()
for(x in monkeynames){
  timespent <- sum(monkeysub[which(monkeysub$Individual == x), 1])
  IndivSumTimes <- c(IndivSumTimes, timespent)
}
SumTimes <- data.frame(monkeynames, IndivSumTimes)

plot(as.factor(monkeynames), as.numeric(IndivSumTimes), data = SumTimes) 
sd(IndivSumTimes)
print("sd for Individual Times: 868.6295")

## plot for habitat vs sum time
HabitatSumTimes <- c()
for(x in habitats){
  timespent <- sum(monkeysub[which(monkeysub$Habitat == x), 1])
  HabitatSumTimes <- c(HabitatSumTimes, timespent)
}
HabSumTimes <- data.frame(habitats, HabitatSumTimes)

plot(as.factor(habitats), as.numeric(HabitatSumTimes), data = HabSumTimes)

## plot for behavior vs sum time
BehaviorSumTimes <- c()
for(x in behaviors){
  timespent <- sum(monkeysub[which(monkeysub$Behavior == x), 1])
  BehaviorSumTimes <- c(BehaviorSumTimes, timespent)
}
BevSumTimes <- data.frame(behaviors, BehaviorSumTimes)

plot(as.factor(behaviors), as.numeric(BehaviorSumTimes), data = BevSumTimes)

## plot for focal ID vs sum time but i cant remember what focal ID actually is
focalid2 <- c(1:61)
focalIDSumTimes <- c()
for(i in 1:61){
  timespent <- sum(monkeysub[which(monkeysub$FocalID == i), 1])
  focalIDSumTimes <- c(focalIDSumTimes, timespent)
}
FocalIDSumTimes <- data.frame(focalid2, focalIDSumTimes)

plot(as.numeric(focalid2), as.numeric(focalIDSumTimes), data = FocalIDSumTimes)
```

## III.2 Behavior Trajectory Plot Attempt (failed...)
```{r eval=FALSE, include=FALSE}
## Behavioral Trajectory Plot (it didnt work)

onefocal <- subset(monkeysub, monkeysub$FocalID == 1.0)

plot(x = as.numeric(onefocal$TimeStamp), y = as.factor(onefocal$Behavior))

## issue: how to treat the time stamp? its not numeric? but it wasnt working as a factor?


## take 2 I found a package??? didnt really look too much into it so hopefully its not sus

library(ggplot2)
library(lubridate)
onefocal$TimeStamp <- as.POSIXct(as.character(onefocal$TimeStamp), format="%X")

ggplot(data = onefocal, aes(x=TimeStamp)) + 
  geom_point(aes(y=Behavior, col=1)) +
  ylab("Behavior") + xlab("Time")


## take 3 on behavior trajectory plot
cumsum <- cumsum(onefocal$Time)
cumsum <- data.frame(cumsum)
onefocal <- cbind(onefocal, cumsum)

ggplot(data = onefocal, aes(x=cumsum)) + 
  geom_point(aes(y=Behavior, col=1)) +
  geom_path(aes(y=Behavior, col=1)) +
  ylab("Behavior") + xlab("Time")

## things that didnt work: geom_segment (needs specific start/end)
## geom_path --> gave the same thing as geom_line
## geom_jitter + geom_path was a wack combo
## geom point + geom_path was similar but still not right

## ok so the issue is that i cant get the lines to cross to a different behavior
## putting a pause in that to try to overlap some plots with just the points

onefocal2 <- subset(monkeysub, monkeysub$FocalID == 2.0)
cumsum2 <- cumsum(onefocal2$Time)
cumsum2 <- data.frame(cumsum2)
onefocal2 <- cbind(onefocal2, cumsum2)

ggplot(data = onefocal2, aes(x=cumsum2)) + 
  geom_point(aes(y=Behavior, col=1)) +
  ylab("Behavior") + xlab("Time")

ggplot(data = onefocal2, aes(x=cumsum2)) + 
  geom_point(aes(y=Behavior, col=1)) +
  geom_point(data = onefocal, aes(x = cumsum, y = Behavior)) +
  ylab("Behavior") + xlab("Time")

## add another one
onefocal3 <- subset(monkeysub, monkeysub$FocalID == 3.0)
cumsum3 <- cumsum(onefocal3$Time)
cumsum3 <- data.frame(cumsum3)
onefocal3 <- cbind(onefocal3, cumsum3)

ggplot(data = onefocal3, aes(x=cumsum3)) + 
  geom_point(aes(y=Behavior, col=1)) +
  ylab("Behavior") + xlab("Time")

ggplot(data = onefocal2, aes(x=cumsum2)) + 
  geom_point(aes(y=Behavior, col=1)) +
  geom_point(data = onefocal, aes(x = cumsum, y = Behavior)) +
  geom_point(data = onefocal3, aes(x = cumsum3, y = Behavior)) +
  ylab("Behavior") + xlab("Time")

onefocal4 <- subset(monkeysub, monkeysub$FocalID == 4.0)
cumsum4 <- cumsum(onefocal4$Time)
cumsum4 <- data.frame(cumsum4)
onefocal4 <- cbind(onefocal4, cumsum4)

onefocal5 <- subset(monkeysub, monkeysub$FocalID == 5.0)
cumsum5 <- cumsum(onefocal5$Time)
cumsum5 <- data.frame(cumsum5)
onefocal5 <- cbind(onefocal5, cumsum5)

```

# IV. Transition Matrices

## IV.1 trying packages bc i didnt want to do it myself
```{r eval=FALSE, include=FALSE}
# new packages
## trying a new package: markovchain
## https://cran.r-project.org/web/packages/markovchain/markovchain.pdf
## its only for discrete time but we're gonna roll with it
library(markovchain)
## prop.table (base R)
## https://www.r-bloggers.com/2021/11/markov-transition-animated-plots/
## this is a lit animated plot but i need the matrix first
## should probably try to add a color scale to that or something  



library(dplyr)
library(tidyr)
library(igraph)

df <- data.frame(monkeysub$Behavior, monkeysub$Time)
transmat <- graph_from_data_frame(df, directed = TRUE)
plot(transmat,edge.arrow.size=.5, vertex.color="pink", vertex.size=3, 
     vertex.frame.color="gray", vertex.label.color="lightblue", 
     vertex.label.cex=.5, vertex.label.dist=2, edge.curved=0.5,layout=layout_with_lgl)
## yeah that didnt do what I wanted it to do nm


## most packages assume that you already have a transition matrix which is a problem


## migrate package
library(migrate)
df2 <- data.frame(monkeysub$Behavior, monkeysub$TimeStamp)
## ok the migrate function didnt work either
## i'm going insane theres no way that I just have to do this manually for all of them??

```

# V. Reset/where I switched to a rmd file
- I tried a whole bunch of stuff trying to get R to automatically give me a transition matrix for a focal and it was not cooperating with me :(
- I went into excel to clean up the csv file a little to make it easier to work with
- Q: is PAUSE a pause on the monkey or the observer
- Q: what does it mean when its the same behavior right on top of another (i've been assuming theres no gap)

```{r}
## data import
monkeyprelimdata <- read.csv("~/Desktop/monkey/monkeyprelimdata.csv")
behaviorshortlist <- c("move", "vig", "feed", "rest", "affil", "agg", "groom", "survey", "PAUSE", "human", "forage", "sleep", "play")
```

## V.1 Isolate focal function
```{r}
## function to isolate a single focal & add a cumulative time column
isolate.focal <- function(i){
  onefocal.i <- subset(monkeyprelimdata, monkeyprelimdata$FocalID == i)
  cumsum.i <- data.frame(cumsum(onefocal.i$TimeSpent))
  onefocal.i <- cbind(onefocal.i, cumsum.i)
  return(onefocal.i)
}
```

# VI. Looking at Individual Focals

## VI.1 Comparing individuals in forest vs road
```{r}
## isolate selected focals (1 forest, 1 road per individual)
## can't say i chose these randomly but like I chose them on vibes of liking the focal numbers lolß

## Caca
focal1 <- isolate.focal(1)
focal43 <- isolate.focal(43)


## afwan
focal35 <- isolate.focal(35)
focal40 <- isolate.focal(40)

## Gado
focal21 <- isolate.focal(21)
focal28 <- isolate.focal(28)
```

```{r}
## create transition matrices

## CACA (1, 43)
trans1 <- matrix(c(0.8421, 0.1053, 0.0526, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0.0102, 0.9898, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0.0028, 0, 0.9972, 0, 0, 0, 0, 0, 0, 0, 0, 0 ,0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0), nrow = 13, ncol = 13)

trans43 <- matrix(c(0.989, 0, 0, 0, 0, 0, 0.0074, 0, 0, 0, 0.0037, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0 ,0,
                   0.0119, 0, 0, 0.9881, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0.006, 0, 0, 0.012, 0, 0, 0.9819, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0.25, 0, 0, 0, 0.75, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0), nrow = 13, ncol = 13)

par(mfrow = c(1, 2))
image(trans1, col = hcl.colors(13, "set 3"))  ## caca jalan
image(trans43, col = hcl.colors(13, "set 3")) ## caca hutan
```

```{r}
## AFWAN (35, 40)

trans35 <- matrix(c(0.9167, 0, 0.0417, 0, 0.0083, 0.0083, 0.025, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0.0154, 0, 0.9730, 0, 0, 0, 0.0039, 0, 0, 0.0077, 0, 0 ,0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0.0105, 0, 0, 0, 0.9895, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0.667, 0, 0, 0, 0.333, 0, 0, 0,
                   0.0606, 0, 0, 0, 0, 0, 0.9394, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0.0789, 0, 0, 0, 0, 0, 0, 0.9211, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0.25, 0, 0, 0, 0.75, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0), nrow = 13, ncol = 13)

trans40 <- matrix(c(0.9798, 0, 0, 0, 0, 0, 0.0101, 0, 0, 0, 0.0101, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0 ,0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0.0021, 0, 0, 0, 0.9958, 0, 0, 0.0021, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0.5, 0, 0, 0, 0, 0, 0.5, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0.1111, 0, 0, 0.8889, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0.1111, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0.8889, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0), nrow = 13, ncol = 13)


par(mfrow = c(1, 2))
image(trans35, col = hcl.colors(13, "set 3")) ## afwan Jalan
image(trans40, col = hcl.colors(13, "set 3")) ## afwan hutan
```

observations
\
-feed to move is such a cycle
\
-human directed behavior is always followed by feed (which makes sense)
\
-just a lot more going on when they are by the road i think
\
\
still struggling to get the axis labels, see behaviorshortlist for the order on the axes

```{r}
## GADO (21, 28)
trans21 <- matrix(c(0.9583, 0, 0, 0.0208, 0, 0, 0, 0, 0, 0.0208, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0 ,0,
                   0, 0, 0, 0.9623, 0, 0, 0, 0, 0, 0.0377, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0.9286, 0, 0, 0, 0.0714, 0, 0, 0,
                   0.0131, 0, 0, 0.0065, 0, 0, 0.9673, 0, 0, 0.0131, 0, 0, 0,
                   0, 0, 0, 0, 0, 0.0069, 0, 0.9931, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0.0118, 0, 0, 0.0296, 0.0118, 0, 0.9467, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0), nrow = 13, ncol = 13)

trans28 <- matrix(c(0.9444, 0, 0.0556, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0.96, 0.04, 0, 0, 0, 0, 0, 0, 0, 0 ,0,
                   0, 0, 0, 0.9722, 0.0278, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0.005, 0.995, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                   0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0), nrow = 13, ncol = 13)


par(mfrow = c(1, 2))
image(trans21, col = hcl.colors(13, "Greens3", rev = 1)) ## gado Jalan
image(trans28, col = hcl.colors(13, "Greens3", rev = 1)) ## gado hutan
```
-wack: human directed not followed by feed (actually check the next, yeah that didnt tell me anything, i was looking to see if maybe it was some other interaction with a human other than being fed)
\
-really cant see the transitions in this one, did a lot of staying still
\
-def feel like the diff monkeys have diff vibes


For next time (after break):
-make a toy data set to try to get it to make a transition matrix
  -double loop
-MarkovChainFit
-sparse representation



# VII. Transition Matrix Function!

make a function to get a transition matrix from a focal

```{r}
behaviorshortlist <-c("Move", "Vigilance", "Feed", "Rest",
               "Affiliative", "Aggressive", "SelfGroom",
               "Survey", "PAUSE", "Human Directed",
               "Forage", "Sleep", "Play")

matrixfunct <- function(dataframe, behaviorshortlist){
  Matrix <- matrix(NA, nrow = length(behaviorshortlist), ncol = length(behaviorshortlist))
  rownames(Matrix) <- colnames(Matrix) <- behaviorshortlist
  for(b1 in behaviorshortlist){
    denomentator <- sum(dataframe$TimeSpent[dataframe$Behavior == b1]) - 
      as.numeric(tail(dataframe$Behavior, 1) == b1) ## accounting for if the last behavior in focal is b1 (doesn't count as a transition opportunity), overall total possible transition opportunities
    b1_ind <- which(dataframe$Behavior == b1)
    for(b2 in behaviorshortlist){
      if(b1 == b2){ ## if staying in the same behavior
        numerator <- denomentator - sum(dataframe$Behavior == b1) + 
          as.numeric(tail(dataframe$Behavior, 1) == b1) 
        ## removing the times they switched to a diff behavior & adding back in that situation if the last behavior is b1
        } else { ## if moving to a different behavior
          numerator <- sum(dataframe$Behavior[b1_ind + 1] == b2, na.rm = T)
          }
      Matrix[b1,b2] <- numerator/denomentator
    }
  }
return(Matrix)
}

Matrix <- matrixfunct(focal28, behaviorshortlist)
rowSums(Matrix)
Matrix
```

problems:
* rows are not adding to 1
  - probably an issue with the tail case (the last behavior)
  - problem solved, it was a typo :)
* still a problem: double behavior thing
  - i think its probably easier to just fix that in the data frame and not try to change the function
  


Up Next (end of January):
* reorganize this file to make sense
* start r package for that function
* make repo (add Henry)
* uncertainty quantification (intervals)
  - predictors/regression?
* new function
  - input: observations, relevant parameters
  - ouput: single scalar likelihood quantile
  - get log like-likelihood 
  - get behaviors

# VIII. Making New Functions

## VIII.1 Pseudo-code mapping

- get log-likelihood function (the spacing on this will probably look wack when i knit but thats life)
_______________________________________________________
P = the probability transition matrix?

get_LL --> function(behaviors, P){
  .
  .
  .
  return(log-likelihood)
}
Check:
If you’ve made behaviors w/particular P, does it give you close to the highest possible value with the first function
1. Simulate behaviors from P
2. Check get_LL is maximized close to P
Then optim function will help us with maximization
^^ alright that is the outline Henry gave 
_________________________________________________________
ok idk what im doing here
ive potentially lost the plot with what this function is supposed to accomplish 

so we want the likelhood --> probability of being in any particular behavior
so this function needs to take in data and calculate probabilty of being in each of the behaviors
that is not matching up with what henry gave above
is the output a vector? like a scalar value for each behavior and then the entire vector adds to 1?

are we assuming we have enough data to make this a normal distribution? is that a bold assumption it feels bold but then we can use dnorm/rnorm???

optim() function for optimization (so like do the derivative part for me once i have the equation?)

________________________________________________________________________
get_LL --> function(behaviors, data){
  n <- length(behaviors) ## number of observations? idk if behaviors is what to put there
  P <- matrixfunct(data, behaviors) ## so this will give us the transition matrix, only works for 1 focal rn
  
  liklihood_eq <- (product (of each entry of the matrix^total number of transitions?))
    ## no thats not right bc it will just give a single number? 
    ## do i do this for each behavior? should this be in the for loop
 
  log-likelihood <- define an empty matrix here with length(behaviors) to add to in the for loop
  
  for(b in behaviors){
    optim(log(liklihood_eq))
    and then add it to the matrix in the right spot in log-likelihood
  }
  return(log-likelihood)
}
___________________________________________________________________________
get_behaviors —> function(P, initial_behavior){
	.
	.
	.
	return(Behaviors)}
	
______________________________
testing that my git sync worked correctly (finally.....legit many hours later)