---
title: "Using depmixS4"
author: "Josephine Ward"
date: "2025-03-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
____________________________________________________________________
# 0. ToC
* I. Setup
* II. Notes
* III. Application Attempts

____________________________________________________________________
# I. Setup
```{r}
## setup
install.packages("depmixS4") ## note there are a bunch of dependent packages
library(depmixS4)

data <- monkeyprelimdata <- read.csv("~/Desktop/MonkeyProject/monkeyprelimdata.csv")

y <- behaviorshortlist <-c("Move", "Vigilance", "Feed", "Rest",
               "Affiliative", "Aggressive", "SelfGroom",
               "Survey", "Human Directed",
               "Forage", "Sleep", "Play") 
```


____________________________________________________________________
# II. Notes
Notes: 
“Dependent mixture models”

- Package goals relevant to us
    - Estimate parameter subject to general linear (in)equality constraints;
    - Fit transition models w/covariates (time-dependent transition matrices)
    - Include covariates in the prior or initial state probabilities 
- Longitudinal/time series data
- Expectation-maximization (EM) algorithm
- 2 steps for using this package
    - 1. Specify model with function “depmix”
    - 2. model fitting w/function “fit”
    - Separated bc good place to check work for time-consuming things
- Functions we may want to use:
    - “depmix” defines the model
        - Then just calling fit on that model should give us the log likelihood 
        - Summary should give the initial state and the transition matrix?
            - Note check if the transition matrix matches

____________________________________________________________________
# III. Application Attempts

```{r}
## gonna run it with their data first to test it

mod.ex <- depmix(response = rt ~ 1, data = speed, nstates = 2, trstart = runif(4))
fm.ex <- fit(mod.ex, emc = em.control(rand = FALSE))
summary(fm.ex)

## that worked as expected nice :)

## apply to our data
set.seed(2025)
# model <- depmix(response = Behavior ~ 1, data = data, nstates = 12, 
#                trstart = runif(12))
## so that didnt work and I think it may be an issue bc our reponse is categorical?

## error in "transInit which i think is part of the response variable, specifically the pstart = 12

fit <- fit(model, emc=em.control(rand=FALSE))
summary(fit)
```

